---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rrtm

<!-- badges: start -->
<!-- badges: end -->

The goal of `rrtm` is to provide efficient R implementations of radiative transfer models useful for vegetation remote sensing.

## Installation

`rrtm` is not currently on CRAN, but
you can install the development version of `rrtm` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ashiklom/rrtm")
```

## Basic usage

This is a basic example of running the PROSPECT 5 leaf radiative transfer model.

```{r basic}
library(rrtm)
out <- prospect5(1.4, 40, 10, 0.01, 0.01)
dim(out)
plot(c(400, 2500), c(0, 1), type = "n",
     xlab = "Wavelength (nm)",
     ylab = "Spectra")
lines(400:2500, out[,,1], col = "red")
lines(400:2500, 1 - out[,,2], col = "blue")
legend("top", legend = c("reflectance", "1 - transmittance"),
       lty = 1, col = c("red", "blue"))
```

Note that the returned object has 3 dimensions, not 2.
This reflects a unique feature of this package: Support for multi-dimensional radiative transfer calculations.
All of the `prospectX` functions can take vectors as inputs and will return multi-dimensional arrays as outputs.

```{r multi}
out2 <- prospect4(1.4, c(30, 40, 50), c(0.01, 0.005, 0.002), 0.01)
plot(c(400, 2500), c(0, 1), type = 'n',
     xlab = "Wavelength (nm)",
     ylab = "Spectra")
matplot(400:2500, f[,,1], type = "l", add = TRUE)
matplot(400:2500, 1 - f[,,2], type = "l", add = TRUE)
```
